### 2019-08-16 EJSSubmission/WeightedSign_EJS_Introduction_02_24_19_1.tex
revision sent to kodu da

Specifically, we propose using product of the generalized sign function and a weight function derived as a transformation of a data-depth function \citep{ref:AoS00461_ZuoSerfling, ref:DIMACS061_Serfling}. Like data-depth functions, the weight functions used in this paper are non-negative reals defined over $\cX \times \cF$, where $\cF$ is a fixed family of probability measures. For every choice of parameters $\mu \in \cX$ and $\BF \in \cF$, in this paper
%
\begin{align}
R (X_{i}; \mu, \BF) = \mu +  S( X_{i}, \mu) W( X_{i}, \BF).
\label{eqn:rank}
\end{align}
%
is used as a robust surrogate for observation $X_{i}$.
Notice that for the trivial choice $W( x, \BF) = | x - \mu_{\BF} |$, 
 $\mu = \mu_{\BF}$,  we get $R (X_{i}; \mu, \BF) = X_{i}$, the original observations. 
 With the other trivial choice of  $W( x, \BF) \equiv 1$, we get the 
 generalized  sign $R (X_{i}; \mu, \BF) = S(X_{i}, \mu) = S_{i}$. 
 However, in this paper we illustrate how using other weight functions
 can lead to interesting robustness and efficiency trade-offs in a variety of situations. 
 The various technical conditions and assumptions that we impose on the 
 weight function $W( x, \BF)$ are valid for weights derived 
 from three well-known data depth functions: the \textit{half-space depth}, 
 the \textit{Mahalanobis depth}, and the \textit{projection depth}.

 %\begin{comment}
\begin{figure}[t]
%\captionsetup{justification=centering, font=footnotesize}
\begin{center}
\begin{tabular}{ll}
%%% These ones have sample size = 700, just right  for EJS
\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_Scatter_700_02_27_19} &
\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_MD_700_02_27_19} \\
\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_TD_700_02_27_19} &
\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_PD_700_02_27_19} \\

@@

%%% These ones have sample size = 1e3, too large file for EJS
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_Scatter_02_24_19} &
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_MD_02_24_19} \\
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_TD_02_24_19} &
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_PD_02_24_19} \\
\end{tabular}
\caption{
An illustrative bivariate scatter plot in the top left panel where the outliers are 
identified with red circles, 
and generalized signs from the same data (black points on the 
unit radius circle, outliers are red points) in the other panels. 
In the top right (bottom left, bottom right) 
panel, weighted signs from the same data with weights obtained using Mahalanobis depth 
(Tukey depth, projection depth respectively) are presented as green triangles 
(outliers are identified by blue triangles). 
}
\label{fig:Fig1}
\end{center}
\end{figure}
%\end{comment}

@@

There are two unknown quantities in the generalized sign function as defined in \eqref{eq:GSign}: $\mu$ and $\BF$, To estimate dispersion and its eigen-structure robustly, we must start with a robust estimator for $\mu$. In Section~\ref{Sec:WSQuantiles} we present the case for \textit{weighted spatial quantiles}, which can be defined and studied in very general spaces 
$\cX$. One special case of this is the {\it weighted spatial median}. As a location estimator, it has several interesting robustness properties and can be shown to be more efficient that some existing robust location estimators, thus making it a perfect candidate to estimate $\mu$. On the other hand, setting $\BF$ as $\BF_X$, i.e. the distribution of $X$ has the clear interpretation of differentially weighting observations based on their inlyingness to the overall data distribution. This mirrors the context of how data depth has been used in the literature (refs). Due to this reason, we assume $\BF \equiv \BF_X$ for the rest of the paper. As and when required, we shall use sample versions of these quantities, stating the theoretical conditions for the corresponding approximation results to hold.

@@

In the rest of this paper, all finite-dimensional vectors are 
column vectors, and for a vector or matrix $a$, the notation $a^{T}$ stands for 
its transpose. The Gaussian distribution with mean $\mu$ and variance $\Sigma$ is denoted
by $N (\mu, \Sigma)$. The identity matrix  is denoted by $\BI$, with or without a 
subscript to denote its dimension. The notations $A^{-1}$ and $det (A)$ respectively 
stand for the inverse and determinant of matrix $A$, whenever these are well-defined.

>>>>

Specifically, we propose using product of the generalized sign function and a weight function derived as a transformation of a data-depth function \citep{ref:DIMACS061_Serfling, ref:AoS00461_ZuoSerfling}. Like data-depth functions, the weight functions used in this paper are non-negative reals defined over $\cX \times \cF$, where $\cF$ is a fixed family of probability measures. For every choice of parameters $\mu \in \cX$ and $\BF \in \cF$, in this paper
%
\begin{align}
R (X_{i}; \mu, \BF) = \mu +  S( X_{i}, \mu) W( X_{i}, \BF).
\label{eqn:rank}
\end{align}
%
is used as a robust surrogate for observation $X_{i}$.
Notice that for the trivial choice $W( x, \BF) = | x - \mu_{\BF} |$, 
 $\mu = \mu_{\BF}$,  we get $R (X_{i}; \mu, \BF) = X_{i}$, the original observations. 
 With the other trivial choice of  $W( x, \BF) \equiv 1$, we get the 
 generalized  sign $R (X_{i}; \mu, \BF) = S(X_{i}, \mu) = S_{i}$. 
 However, in this paper we illustrate how using other weight functions
 can lead to interesting robustness and efficiency trade-offs in a variety of situations. 
 The various technical conditions and assumptions that we impose on the 
 weight function $W( x, \BF)$ are valid for weights derived 
 from three well-known data depth functions: the \textit{half-space depth}, 
 the \textit{Mahalanobis depth}, and the \textit{projection depth}.

 %\begin{comment}
\begin{figure}[t]
%\captionsetup{justification=centering, font=footnotesize}
\begin{center}
%\begin{tabular}{ll}
%%% These ones have sample size = 700, just right  for EJS
\includegraphics[width=\textwidth]{./Plots_SC/WeightedSignAll}
%\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_Scatter_700_02_27_19} &
%\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_MD_700_02_27_19} \\
%\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_TD_700_02_27_19} &
%\includegraphics[width=0.45\textwidth]{./Plots_SC/WeightedSign_PD_700_02_27_19} \\

@@

%%% These ones have sample size = 1e3, too large file for EJS
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_Scatter_02_24_19} &
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_MD_02_24_19} \\
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_TD_02_24_19} &
%\includegraphics[width=0.45\textwidth]{../Plots_SC/WeightedSign_PD_02_24_19} \\
%\end{tabular}
\caption{
An illustrative bivariate scatter plot in the top left panel where the outliers are 
identified with red circles, 
and generalized signs from the same data (black points on the 
unit radius circle, outliers are red points) in the other panels. 
In the top right (bottom left, bottom right) 
panel, weighted signs from the same data with weights obtained using Mahalanobis depth 
(Tukey depth, projection depth respectively) are presented as green triangles 
(outliers are identified by blue triangles). 
}
\label{fig:Fig1}
\end{center}
\end{figure}
%\end{comment}

@@

There are two unknown quantities in the generalized sign function as defined in \eqref{eq:GSign}: $\mu$ and $\BF$, To estimate dispersion and its eigen-structure robustly, we must start with a robust estimator for $\mu$. In Section~\ref{Sec:WSQuantiles} we present the case for \textit{weighted spatial quantiles}, which can be defined and studied in very general spaces 
$\cX$. One special case of this is the {\it weighted spatial median}. As a location estimator, it has several interesting robustness properties and can be shown to be more efficient that some existing robust location estimators, thus making it a perfect candidate to estimate $\mu$. On the other hand, setting $\BF$ as $\BF_X$, i.e. the distribution of $X$ has the clear interpretation of differentially weighting observations based on their inlyingness to the overall data distribution. This mirrors the context of how data depth has been used in the literature \cite{LiuPareliusSingh99, ref:DIMACS061_Serfling}. Due to this reason, we assume $\BF \equiv \BF_X$ for the rest of the paper. As and when required, we shall use sample versions of these quantities, stating the theoretical conditions for the corresponding approximation results to hold.

@@

In the rest of this paper, all finite-dimensional vectors are column vectors, and for a vector or matrix $a$, the notation $a^{T}$ stands for its transpose. The Gaussian distribution with mean $\mu$ and variance $\Sigma$ is denoted by $N (\mu, \Sigma)$. The identity matrix  is denoted by $\BI$, with or without a subscript to denote its dimension. The notations $A^{-1}, det (A), \lambda_{\min} (A), \lambda_{\max} (A)$ respectively stand for the inverse, determinant, minimum and maximum eigenvalues of a matrix $A$, whenever these are well-defined. For a scalar or vector valued random variable $Y$, $\BE Y$ denotes its expected value, while $\BV Y$ denotes its variance or covariance matrix.

### 2019-08-16 EJSSubmission/WeightedSign_EJS_Simulations_02_24_19_1.tex
revision sent to kodu da

We report numerical simulation studies on several properties relating to 
$\tilde{\Sigma}$ and  $\Sigma_{*}$, and their eigenvalues and eigenvectors, 
on datasets with or without influential points, to illustrate the finite sample 
efficiency and robustness properties of the proposed weighted estimators. We compare 
these proposed estimators with techniques that exists in literature, specifically, 
the Sign Covariance Matrix (SCM) and Tyler's shape matrix \citep{ref:AoS87234_Tyler}.

>>>>

We report a number of numerical simulation studies on several properties relating to 
$\tilde{\Sigma}$ and  $\Sigma_{*}$, and their eigenvalues and eigenvectors, 
on datasets with or without influential points, to illustrate the finite sample 
efficiency and robustness properties of the proposed weighted estimators. We compare 
these proposed estimators with techniques that exists in literature, specifically, 
the Sign Covariance Matrix (SCM) and Tyler's shape matrix \citep{ref:AoS87234_Tyler}.

### 2019-08-16 EJSSubmission/WeightedSign_EJS_WeightedSpatialQuantiles_02_24_19_1.tex
revision sent to kodu da

\section{Weighted Spatial Quantiles}
\label{Sec:WSQuantiles}

@@

\begin{description}
\item[A1:] $\Psi (q; 0, \BF)$ is finite for all $q \in \cX \subseteq \BR^{p}$ 
and has a unique minimizer $q_{0}$.

 \item[A2:] $\Psi (q; 0, \BF)$ is twice differentiable at $q_{0}$ and the second 
 derivative is positive definite.

 \item[A3:] $\BE W^{2} (X, \BF)  S (X; q) S^{T} (X; q)$ exists for all 
 $q$ in a neighborhood of $q_{0}$. 
 \end{description}
 These assumptions are very broad-based and general. The first one essentially requires 
 the existence of a population parameter, the second one requires that the minimization 
 approach is meaningful in the population, and the third one essentially requires that the 
 weight function has a finite variance. No further restrictions are placed on the tuning parameter $\BF$ or the choice of the weight function.

@@

\subsection{Asymptotic efficiency of weighted spatial median}
Let $V_{W} = \Psi_{2W}^{-1} \Psi_{1W} \Psi_{2W}^{-1}$ be the asymptotic 
variance of \hl{$\hat{q}_{n,W}$} from Theorem~\ref{Thm:WSMedian}, where we use the 
subscript ``${}_{W}$'' to denote that this depends on the weight function. 
We use the notation 
$V_{1}$ for the case where $W (x, \BF) \equiv 1$, that is, all weights are one.
The asymptotic relative efficiency of two statistics is the $p$-th root of 
the reciprocals of their determinants. That is, 
\ban 
ARE (\hat{q}_{nW}, \hat{q}_{n}) = \Bigl\{\frac{det(V_{1})}{det(V_{W})} \Bigr\}^{1/p}.
\ean
One easy result from Theorem~\ref{Thm:WSMedian} is that under reasonable conditions the
asymptotic relative efficiency of the weighted spatial median over the unweighted version
is always greater than one. We document this in the following corollary:

\begin{Corollary}
\label{Cor:ARE}
Assume that the weight function $W (X, \BF)$ is bounded above by $k$ for some $k > 0$, and the matrices $\Psi_1 = \BE S(X; q_0) S^T(X; q_0)$ and $\Psi_{1W}$ are positive definite. Then
%
\ban
ARE (\hat{q}_{nW}, \hat{q}_{n}) \geq
\frac{ \lambda_{\min} (\Psi_1) \lambda_{\min}^2 (\Psi_{2W})}
{k \lambda_{\max} (\Psi_1) \lambda_{\max}^2 (\Psi_2)}
\ean
%
Consequently, if $k < r \equiv \lambda_{\min} (\Psi_1) \lambda_{\min}^2 (\Psi_{2W})/(\lambda_{\max} (\Psi_1) \lambda_{\max}^2 (\Psi_2))$ then this asymptotic relative efficiency is larger than 1. 
\end{Corollary}

@@

There are several alternative formulations, see the above reference and citations within it for details. The spherically symmetric case is realized with $\Sigma = \sigma^{2} \BI_{p}$ for some 
$\sigma^{2} > 0$. 
We fix the notation $Z = \Sigma^{-1/2} (X - \mu)$, and let 
$Z \sim \BF_{Z}$. Note that $\BF_{Z}$ is a spherically symmetric distribution and hence
depends only on $|z|$, and  $\BE Z = \mathbf{0}_{p} \in \BR^{p}$ and $\BV Z = \BI_{p}$.

\begin{Corollary}
\label{Cor:Elliptic_ARE}
Assume that $\BF \equiv \BF_X$ is an elliptically symmetric distribution, with location parameter $\mu = q_{0}$, $\Sigma$ positive definite. Also assume that the weight function is affine invariant, that is:
\ban
W (A x + b, A \BF + b) = W (x, \BF),
\ean
Then  the weight function satisfies $W (X, \BF) = W (Z, \BF_{Z})$, and hence may be further restricted to be a function of $|Z|$. Then $S (X; q_0)$ is uniform on the $\sqrt c$-radius ball 
in $\BR^{p}$ centered at the origin, and is independent of $|X - q_{0}|$, while $W (X, \BF)$ is a function of  $|X - q_{0}|$. That is, the conditions of Corollary~\ref{Cor:ARE} are satisfied. 
\end{Corollary}

Obtaining affine invariant weight functions is not challenging, in fact, most 
such functions arising in the context of data-depths are affine invariant. 
%Corollary~\ref{Cor:Elliptic_ARE} implies that there is a wide frame of distributions and  choices of weight functions where there is a benefit to considering weighted spatial quantiles.
In actual computations, in place of $W (x, \BF)$ we propose 
using $W (x, \BF_{n})$ where
$ \BF_{n}(z) = (np)^{-1} \sum_{i = 1}^{n} \sum_{j=1}^p
\cI (X_{ij} \leq z_j ), z \in \BR^p
$ is the empirical distribution function. Up to first order asymptotics, the analysis remain unchanged from the above for this modified weight function.

@@

It is easy to show that results in this paper are valid for the weight functions
($i$) $W_{HSD} (X) = \BF_{Z} (| Z |)$ derived from the \textit{half-space depth},  
($ii$) $W_{MhD} (X) = |Z|^{2}/(1 + | Z |^2)$ derived from the \textit{Mahalanobis depth}, 
and  
($iii$) $W_{PD} (X) = |Z|/(1+ | Z |/MAD(\BF_{Z}))$, where $MAD$ stands for median 
absolute deviation, derived from the \textit{projection depth}. We omit the technical 
details. These three weight functions give a center-inward partial ordering, thus essentially 
quantifying \textit{peripherality} instead of \textit{depth}.
Note however, that our results below are of much more general form, and 
these three special choices of weights only serve as important illustrative examples 
to achieve desirable robustness and efficiency balance in data analysis.

>>>>

\section{The Weighted Spatial Median}
\label{Sec:WSQuantiles}

@@

\begin{description}
\item[(A1)] $\Psi (q; 0, \BF)$ is finite for all $q \in \cX \subseteq \BR^{p}$ 
and has a unique minimizer $q_{0}$.

\item[(A2)] $\Psi (q; 0, \BF)$ is twice differentiable at $q_{0}$ and the second derivative is positive definite.

\item[(A3)] $\BE W^{2} (X, \BF)  S (X; q) S^{T} (X; q)$ exists for all $q$ in a neighborhood of $q_{0}$. 
\end{description}
 These assumptions are very broad-based and general. The first one essentially requires 
 the existence of a population parameter, the second one requires that the minimization 
 approach is meaningful in the population, and the third one essentially requires that the 
 weight function has a finite variance. No further restrictions are placed on the tuning parameter $\BF$ or the choice of the weight function.

@@

\subsection{Asymptotic efficiency of weighted spatial median}
Let $V_{W} = \Psi_{2W}^{-1} \Psi_{1W} \Psi_{2W}^{-1}$ be the asymptotic 
variance of $\hat{q}_{n,W}$ from Theorem~\ref{Thm:WSMedian}, where we use the 
subscript ``${}_{W}$'' to denote that this depends on the weight function. 
We use the notation 
$V_{1}$ for the case where $W (x, \BF) \equiv 1$, that is, all weights are one.
The asymptotic relative efficiency of two statistics is the $p$-th root of 
the reciprocals of their determinants. That is, 
\ban 
ARE (\hat{q}_{nW}, \hat{q}_{n}) = \Bigl\{\frac{det(V_{1})}{det(V_{W})} \Bigr\}^{1/p}.
\ean
One easy result from Theorem~\ref{Thm:WSMedian} is that under reasonable conditions the
asymptotic relative efficiency of the weighted spatial median over the unweighted version
is always greater than one. We document this in the following corollary:

\begin{Corollary}
\label{Cor:ARE}
Assume that the weight function $W (X, \BF)$ is bounded above by $k$ for some $k > 0$, and the matrices $\Psi_1 = \BE S(X; q_0) S^T(X; q_0)$ and $\Psi_{1W}$ are positive definite. Then
%
\ban
ARE (\hat{q}_{nW}, \hat{q}_{n}) \geq
\frac{ \lambda_{\min} (\Psi_1) \lambda_{\min}^2 (\Psi_{2W})}
{k \lambda_{\max} (\Psi_1) \lambda_{\max}^2 (\Psi_2)}
\ean
%
Consequently, if $k / \lambda_{\min}^2 (\Psi_{2W}) < \lambda_{\min} (\Psi_1) /(\lambda_{\max} (\Psi_1) \lambda_{\max}^2 (\Psi_2))$ then this asymptotic relative efficiency is larger than 1. 
\end{Corollary}

@@

There are several alternative formulations, see the above reference and citations within it for details.

\begin{Corollary}\label{Cor:Elliptic_ARE}
Assume that $\BF \equiv \BF_X$ is an elliptically symmetric distribution, with location parameter $\mu = q_{0}$ and the covariance matrix $\Sigma$ satisfies the following conditions:
%
\begin{enumerate}
\item $Tr(\Sigma^4) = o( Tr^2 (\Sigma^2) )$,

\item $\frac{Tr^4(\Sigma)}{ Tr^2 (\Sigma^2)} \exp \left\{ - \frac{Tr^2 (\Sigma)}
{ 128 \lambda_{\max}^2 (\Sigma)} \right\} = o(1)$,

\item $\exp \left\{ - \frac{Tr^2 (\Sigma)}{ 256 \lambda_{\max}^2 (\Sigma)} \right\} = 
o\left( \min\left( \frac{\lambda_{\max}(\Sigma) }{Tr( \Sigma)},
\frac{\lambda_{\min}(\Sigma)}{\lambda_{\max}(\Sigma)} \right) \right)$,

\item $\lambda_{\max}(\Sigma) = o( Tr(\Sigma))$.
\end{enumerate}

Also assume that the weight function is (a) bounded above by some $W_{\max} > 0$, (b) affine invariant, i.e. $ W (Ax + b, A\BF + b) = W (x, \BF)$ for $A \in \BR^{p \times p}, b \in \BR^p$, and (c) satisfies the following:
%
\begin{equation}\label{eqn:CorElliptic_ARE_eqn}
\frac{W_{\max}}{[ \BE | X - q_0|^{-1} W(X, \BF)]^2 } <
\frac{\lambda_{\min} (\Psi_1)}{\lambda_{\max} (\Psi_1) [ \BE | X - q_0|^{-1}]^2]}.
\end{equation}
%
Then we have $ARE( \hat q_{n,W}, \hat q_n) > 1$.
\end{Corollary}

The conditions 1-4 in Corollary~\ref{Cor:Elliptic_ARE} are due to \cite{ref:JASA151658_WangPengLi}, and ensure that the eigenvalues of $\Sigma$ are bounded away from 0 and $\infty$. Corollary~\ref{Cor:Elliptic_ARE} can be proved using Corollary~\ref{Cor:ARE}, the fact that for elliptical distributions $\Psi_1$ is non-singular \cite{ref:SPL12765_Taskinenetal}, then using and expanding upon Lemma A.5 in \cite{ref:JASA151658_WangPengLi}. We give the technical details in the supplementary material.

Obtaining affine invariant weight functions is not challenging. Most functions arising in the context of data depths are affine invariant. Corollary~\ref{Cor:Elliptic_ARE} implies that there is a wide frame of distributions and  choices of weight functions where there is a benefit to considering weighted spatial medians. In fact, the choice of weight functions is even broader: only location invariance is required for Corollary~\ref{Cor:Elliptic_ARE} to hold. We only choose to restrict ourselves here because the subsequent analysis uses weights based on affine invariant data depth functions. Also note that in case \eqref{eqn:CorElliptic_ARE_eqn} does not hold for the original form of a weight function, one can scale all weights by an appropriate constant to reduce the value in the left hand side of \eqref{eqn:CorElliptic_ARE_eqn} to satisfy the condition.

In actual computations, in place of $W (x, \BF)$ we propose using $W (x, \BF_{n})$ where
$ \BF_{n}(z) = (np)^{-1} \sum_{i = 1}^{n} \sum_{j=1}^p
\cI (X_{ij} \leq z_j ), z \in \BR^p
$ is the empirical distribution function. Up to first order asymptotics, the analysis remain unchanged from the above for this modified weight function.

@@

The spherically symmetric case of an elliptical distribution is realized with $\Sigma = \sigma^{2} \BI_{p}$ for some $\sigma^{2} > 0$. We fix the notation $Z = \Sigma^{-1/2} (X - \mu)$, and let 
$Z \sim \BF_{Z}$. Note that $\BF_{Z}$ is a spherically symmetric distribution and hence
depends only on $|z|$, and $\BE Z = \mathbf{0}_{p} \in \BR^{p}$ and $\BV Z = \BI_{p}$. Taking affine invariant data depth functions as weights ensures that $W(X, \BF) = W(Z, \BF_Z)$. It is now easy to show that results in this paper are valid for the weight functions (with appropriate scaling to satisfy \eqref{eqn:CorElliptic_ARE_eqn})
($i$) $W_{HSD} (X) = \BF_{Z} (| Z |)$ derived from the \textit{half-space depth},  
($ii$) $W_{MhD} (X) = |Z|^{2}/(1 + | Z |^2)$ derived from the \textit{Mahalanobis depth}, 
and ($iii$) $W_{PD} (X) = |Z|/(1+ | Z |/MAD(\BF_{Z}))$, where $MAD$ stands for median 
absolute deviation, derived from the \textit{projection depth}. We omit the technical 
details. These three weight functions give a center-inward partial ordering, thus essentially 
quantifying \textit{peripherality} instead of \textit{depth}.
Note however, that our results below are of much more general form, and 
these three special choices of weights only serve as important illustrative examples 
to achieve desirable robustness and efficiency balance in data analysis.
